---
title: "Final Thesis Analyses: Loggerhead Clutch Relocation at Mon Repos"
author: "Madison Scott"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE}
# ============================================================
# FINAL THESIS ANALYSES
# ============================================================
#Raw data are not publicly available due to a data-sharing agreement with the Queensland Department of the Environment, Tourism, Science and Innovation (DETSI).The repository contains all code required to reproduce analyses and figures. Access to the original data must be requested through DETSI.
#
# This RMarkdown is intended to recreate the final statistical
# tests, summary tables, and figures used in the thesis.
#
# Expected input files:
#   data/qnr_dataset2.csv
#   data/badnest(1).csv
#   data/MR24_emergeEditV1 copy.csv
#
# If these files are in the same folder as this RMarkdown, the
# fallback paths below should still work.
# ============================================================

required_packages <- c(
  "tidyverse", "lubridate", "betareg", "lmtest", "emmeans",
  "lme4", "broom", "broom.mixed", "gt", "patchwork"
)

installed <- rownames(installed.packages())
missing_packages <- setdiff(required_packages, installed)
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

library(tidyverse)
library(lubridate)
library(betareg)
library(lmtest)
library(emmeans)
library(lme4)
library(broom)
library(broom.mixed)
library(gt)
library(patchwork)

theme_set(theme_classic(base_size = 14))

dir.create("figures/final", recursive = TRUE, showWarnings = FALSE)
dir.create("tables/final", recursive = TRUE, showWarnings = FALSE)
dir.create("outputs/model_summaries", recursive = TRUE, showWarnings = FALSE)
dir.create("data/cleaned", recursive = TRUE, showWarnings = FALSE)

# Helper function for beta regression proportions.
beta_transform <- function(y) {
  n <- length(y[!is.na(y)])
  (y * (n - 1) + 0.5) / n
}

find_file <- function(possible_paths) {
  existing <- possible_paths[file.exists(possible_paths)]
  if (length(existing) == 0) {
    stop("Could not find file. Checked: ", paste(possible_paths, collapse = ", "))
  }
  existing[1]
}

qmr_path <- find_file(c("data/qnr_dataset2.csv", "qnr_dataset2.csv", "/mnt/data/qnr_dataset2.csv"))
badnest_path <- find_file(c("data/badnest(1).csv", "badnest(1).csv", "/mnt/data/badnest(1).csv"))
raw_path <- find_file(c("data/MR24_emergeEditV1 copy.csv", "MR24_emergeEditV1 copy.csv", "/mnt/data/MR24_emergeEditV1 copy.csv"))

qmr_dataset2 <- read.csv(qmr_path)
badnest <- read.csv(badnest_path)
raw_emerge <- read.csv(raw_path)
```

# 1. Final data preparation checks

```{r data-prep}
# Standardise final analysis columns.
qmr_dataset2 <- qmr_dataset2 %>%
  mutate(
    RELOC_binary = case_when(
      !is.na(RELOC_binary) ~ as.character(RELOC_binary),
      !is.na(reloc_binary) ~ as.character(reloc_binary),
      is.na(RELOC) ~ "INS",
      TRUE ~ "R"
    ),
    RELOC_binary = case_when(
      RELOC_binary %in% c("INS", "In situ", "In Situ", "In situ ") ~ "INS",
      RELOC_binary %in% c("R", "Relocated", "relocated") ~ "R",
      TRUE ~ RELOC_binary
    ),
    Nest_management = recode(RELOC_binary, "INS" = "In situ", "R" = "Relocated"),
    RELOC_type = case_when(
      RELOC_binary == "INS" ~ "INS",
      TRUE ~ as.character(RELOC_type)
    ),
    RELOC_type = recode(RELOC_type,
                        "INS" = "In situ",
                        "DUNE" = "Dune",
                        "HATCH" = "Hatchery",
                        "CAGE" = "Cage"),
    RELOC_type = factor(RELOC_type, levels = c("In situ", "Dune", "Hatchery", "Cage")),
    Nest_management = factor(Nest_management, levels = c("In situ", "Relocated")),
    CNO = as.numeric(CNO),
    AGECL = factor(AGECL, levels = paste0("A", sort(as.numeric(gsub("A", "", unique(AGECL[!is.na(AGECL)])))))),
    ESUC_prop = ifelse(ESUC > 1, ESUC / 100, ESUC),
    HSUC_prop = ifelse(HSUC > 1, HSUC / 100, HSUC),
    ESUC_prop_beta = beta_transform(ESUC_prop),
    TAG = as.factor(TAG),
    LDATE_date = suppressWarnings(as.Date(LDATE_date)),
    LDATE_numeric = as.numeric(LDATE_numeric)
  ) %>%
  filter(!is.na(ESUC_prop), !is.na(ESUC_prop_beta))

badnest <- badnest %>%
  mutate(
    RELOC_binary = case_when(
      !is.na(RELOC_binary) ~ as.character(RELOC_binary),
      !is.na(reloc_binary) ~ as.character(reloc_binary),
      is.na(RELOC) ~ "INS",
      TRUE ~ "R"
    ),
    RELOC_binary = case_when(
      RELOC_binary %in% c("INS", "In situ", "In Situ", "In situ ") ~ "INS",
      RELOC_binary %in% c("R", "Relocated", "relocated") ~ "R",
      TRUE ~ RELOC_binary
    ),
    Nest_management = recode(RELOC_binary, "INS" = "In situ", "R" = "Relocated"),
    RELOC_type = case_when(
      RELOC_binary == "INS" ~ "INS",
      TRUE ~ as.character(RELOC_type)
    ),
    RELOC_type = recode(RELOC_type,
                        "INS" = "In situ",
                        "DUNE" = "Dune",
                        "HATCH" = "Hatchery",
                        "CAGE" = "Cage"),
    RELOC_type = factor(RELOC_type, levels = c("In situ", "Dune", "Hatchery", "Cage")),
    Nest_management = factor(Nest_management, levels = c("In situ", "Relocated")),
    CNO = as.numeric(CNO),
    AGECL = factor(AGECL, levels = paste0("A", sort(as.numeric(gsub("A", "", unique(AGECL[!is.na(AGECL)])))))),
    ESUC_prop = ifelse(ESUC > 1, ESUC / 100, ESUC),
    HSUC_prop = ifelse(HSUC > 1, HSUC / 100, HSUC),
    ESUC_prop_beta = beta_transform(ESUC_prop),
    TAG = as.factor(TAG),
    LDATE_numeric = as.numeric(LDATE_numeric)
  ) %>%
  filter(!is.na(ESUC_prop), !is.na(ESUC_prop_beta))

stage_cols <- c("UD", "UH2", "UH3", "UH4", "UH5", "UH6")
stage_labels <- c(
  UD = "Undeveloped",
  UH2 = "Stage 2",
  UH3 = "Stage 3",
  UH4 = "Stage 4",
  UH5 = "Stage 5",
  UH6 = "Stage 6"
)

# Save final cleaned datasets
write.csv(qmr_dataset2, "data/cleaned/qmr_dataset2_cleaned.csv", row.names = FALSE)
write.csv(badnest, "data/cleaned/badnest_cleaned.csv", row.names = FALSE)

# Basic counts used in thesis
overall_summary <- qmr_dataset2 %>%
  summarise(
    n_clutches = n(),
    n_individuals = n_distinct(TAG),
    mean_ESUC = mean(ESUC, na.rm = TRUE),
    se_ESUC = sd(ESUC, na.rm = TRUE) / sqrt(sum(!is.na(ESUC)))
  )

reloc_counts <- qmr_dataset2 %>%
  count(Nest_management)

overall_summary
reloc_counts
```

# 2. Relocation status and emergence success

```{r relocation-status-model}
reloc_summary <- qmr_dataset2 %>%
  group_by(Nest_management) %>%
  summarise(
    n = n(),
    mean_ESUC = mean(ESUC, na.rm = TRUE),
    sd_ESUC = sd(ESUC, na.rm = TRUE),
    se_ESUC = sd_ESUC / sqrt(n),
    mean_ESUC_prop = mean(ESUC_prop, na.rm = TRUE),
    se_ESUC_prop = sd(ESUC_prop, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

write.csv(reloc_summary, "tables/final/relocation_status_summary.csv", row.names = FALSE)

beta_reloc <- betareg(ESUC_prop_beta ~ Nest_management, data = qmr_dataset2)
beta_reloc_null <- betareg(ESUC_prop_beta ~ 1, data = qmr_dataset2)

reloc_lrtest <- lmtest::lrtest(beta_reloc, beta_reloc_null)

sink("outputs/model_summaries/beta_relocation_status.txt")
print(summary(beta_reloc))
print(reloc_lrtest)
print(confint(beta_reloc))
sink()

summary(beta_reloc)
reloc_lrtest
confint(beta_reloc)
```

```{r figure-2-relocation-status, fig.width=7, fig.height=5}
fig2 <- ggplot(qmr_dataset2, aes(x = Nest_management, y = ESUC_prop, fill = Nest_management)) +
  geom_boxplot(alpha = 0.75, outlier.shape = NA, width = 0.55) +
  geom_jitter(width = 0.15, alpha = 0.35, size = 1.7) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Nest relocation status",
    y = "Emergence success",
    fill = "Nest management"
  ) +
  theme(legend.position = "none")

fig2
ggsave("figures/final/Figure_2_relocation_status_ESUC.png", fig2, width = 7, height = 5, dpi = 300)
```

# 3. Developmental-stage mortality by relocation status

```{r stage-data-relocation-status}
stage_data_binary <- qmr_dataset2 %>%
  mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
  filter(total_stage_dead > 0, !is.na(Nest_management)) %>%
  mutate(across(all_of(stage_cols), ~ .x / total_stage_dead, .names = "{.col}_prop")) %>%
  pivot_longer(
    cols = ends_with("_prop"),
    names_to = "stage",
    values_to = "prop"
  ) %>%
  mutate(
    stage = gsub("_prop", "", stage),
    stage_label = recode(stage, !!!stage_labels),
    stage_label = factor(stage_label, levels = stage_labels)
  )

stage_summary_binary <- stage_data_binary %>%
  group_by(Nest_management, stage_label) %>%
  summarise(
    n = n(),
    mean_prop = mean(prop, na.rm = TRUE),
    sd = sd(prop, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )

write.csv(stage_summary_binary, "tables/final/stage_mortality_by_relocation_status.csv", row.names = FALSE)

stage_binary_models <- map(stage_cols, function(stage_var) {
  data_stage <- qmr_dataset2 %>%
    mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
    filter(total_stage_dead > 0, !is.na(Nest_management)) %>%
    mutate(
      stage_prop = .data[[stage_var]] / total_stage_dead,
      stage_prop_beta = beta_transform(stage_prop)
    )

  model <- betareg(stage_prop_beta ~ Nest_management, data = data_stage)

  broom::tidy(model) %>%
    filter(component == "mean", term != "(Intercept)") %>%
    mutate(stage = stage_var)
}) %>%
  bind_rows() %>%
  mutate(
    p_holm = p.adjust(p.value, method = "holm"),
    stage_label = recode(stage, !!!stage_labels)
  )

write.csv(stage_binary_models, "tables/final/stage_beta_models_by_relocation_status.csv", row.names = FALSE)

stage_binary_models
```

```{r figure-3-stage-relocation-status, fig.width=8, fig.height=5}
fig3 <- ggplot(stage_summary_binary, aes(x = stage_label, y = mean_prop, color = Nest_management, group = Nest_management)) +
  geom_point(size = 3, position = position_dodge(width = 0.35)) +
  geom_line(position = position_dodge(width = 0.35)) +
  geom_errorbar(aes(ymin = mean_prop - sd, ymax = mean_prop + sd),
                width = 0.15, position = position_dodge(width = 0.35)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Developmental stage at death",
    y = "Mean proportion ± SD",
    color = "Nest management"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

fig3
ggsave("figures/final/Figure_3_stage_mortality_relocation_status.png", fig3, width = 8, height = 5, dpi = 300)
```

# 4. Relocation type and emergence success

```{r relocation-type-model}
reloc_type_summary <- qmr_dataset2 %>%
  filter(!is.na(RELOC_type)) %>%
  group_by(RELOC_type) %>%
  summarise(
    n = n(),
    mean_ESUC = mean(ESUC, na.rm = TRUE),
    sd_ESUC = sd(ESUC, na.rm = TRUE),
    se_ESUC = sd_ESUC / sqrt(n),
    .groups = "drop"
  )

write.csv(reloc_type_summary, "tables/final/relocation_type_summary.csv", row.names = FALSE)

kw_reloc_type <- kruskal.test(ESUC_prop ~ RELOC_type, data = qmr_dataset2)

beta_reloc_type <- betareg(ESUC_prop_beta ~ RELOC_type, data = qmr_dataset2)
beta_reloc_type_null <- betareg(ESUC_prop_beta ~ 1, data = qmr_dataset2)
reloc_type_lrtest <- lmtest::lrtest(beta_reloc_type, beta_reloc_type_null)

reloc_type_emm <- emmeans(beta_reloc_type, ~ RELOC_type, type = "response")
reloc_type_pairs <- pairs(reloc_type_emm, adjust = "BH")

sink("outputs/model_summaries/beta_relocation_type.txt")
print(kw_reloc_type)
print(summary(beta_reloc_type))
print(reloc_type_lrtest)
print(reloc_type_emm)
print(reloc_type_pairs)
sink()

kw_reloc_type
summary(beta_reloc_type)
reloc_type_lrtest
reloc_type_emm
reloc_type_pairs
```

```{r figure-4-relocation-type, fig.width=7, fig.height=5}
fig4 <- qmr_dataset2 %>%
  filter(RELOC_type != "In situ") %>%
  ggplot(aes(x = RELOC_type, y = prop_UH, fill = RELOC_type)) +
  geom_boxplot(alpha = 0.75, outlier.shape = NA, width = 0.55) +
  geom_jitter(width = 0.15, alpha = 0.35, size = 1.7) +
  labs(
    x = "Relocation type",
    y = "Proportion unhatched eggs",
    fill = "Relocation type"
  ) +
  theme(legend.position = "none")

fig4
ggsave("figures/final/Figure_4_unhatched_by_relocation_type.png", fig4, width = 7, height = 5, dpi = 300)
```

# 5. Developmental-stage mortality by relocation type

```{r stage-data-relocation-type}
stage_data_type <- qmr_dataset2 %>%
  mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
  filter(total_stage_dead > 0, !is.na(RELOC_type)) %>%
  mutate(across(all_of(stage_cols), ~ .x / total_stage_dead, .names = "{.col}_prop")) %>%
  pivot_longer(
    cols = ends_with("_prop"),
    names_to = "stage",
    values_to = "prop"
  ) %>%
  mutate(
    stage = gsub("_prop", "", stage),
    stage_label = recode(stage, !!!stage_labels),
    stage_label = factor(stage_label, levels = stage_labels)
  )

stage_summary_type <- stage_data_type %>%
  group_by(RELOC_type, stage_label) %>%
  summarise(
    n = n(),
    mean_prop = mean(prop, na.rm = TRUE),
    sd = sd(prop, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )

write.csv(stage_summary_type, "tables/final/stage_mortality_by_relocation_type.csv", row.names = FALSE)

stage_type_models <- map(stage_cols, function(stage_var) {
  data_stage <- qmr_dataset2 %>%
    mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
    filter(total_stage_dead > 0, !is.na(RELOC_type)) %>%
    mutate(
      stage_prop = .data[[stage_var]] / total_stage_dead,
      stage_prop_beta = beta_transform(stage_prop)
    )

  model <- betareg(stage_prop_beta ~ RELOC_type, data = data_stage)
  emm <- emmeans(model, ~ RELOC_type, type = "response")
  pairwise <- as.data.frame(pairs(emm, adjust = "BH")) %>%
    mutate(stage = stage_var)

  model_tidy <- broom::tidy(model) %>%
    filter(component == "mean", term != "(Intercept)") %>%
    mutate(stage = stage_var)

  list(model = model_tidy, pairwise = pairwise)
})

stage_type_model_table <- map_dfr(stage_type_models, "model") %>%
  mutate(
    p_holm = p.adjust(p.value, method = "holm"),
    stage_label = recode(stage, !!!stage_labels)
  )

stage_type_pairwise_table <- map_dfr(stage_type_models, "pairwise") %>%
  mutate(
    p_holm = p.adjust(p.value, method = "holm"),
    stage_label = recode(stage, !!!stage_labels)
  )

write.csv(stage_type_model_table, "tables/final/stage_beta_models_by_relocation_type.csv", row.names = FALSE)
write.csv(stage_type_pairwise_table, "tables/final/stage_pairwise_by_relocation_type.csv", row.names = FALSE)

stage_type_model_table
stage_type_pairwise_table
```

```{r figure-5-stage-relocation-type, fig.width=10, fig.height=5}
fig5 <- ggplot(stage_data_type, aes(x = stage_label, y = prop, fill = RELOC_type)) +
  geom_boxplot(alpha = 0.75, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(
    x = "Developmental stage at death",
    y = "Proportion of unhatched eggs",
    fill = "Nest management"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

fig5
ggsave("figures/final/Figure_5_stage_mortality_relocation_type.png", fig5, width = 10, height = 5, dpi = 300)
```

# 6. Age class and emergence success

```{r age-class-model}
age_data <- qmr_dataset2 %>%
  filter(!is.na(AGECL), !is.na(ESUC_prop_beta))

age_summary <- age_data %>%
  group_by(AGECL) %>%
  summarise(
    n = n(),
    mean_ESUC = mean(ESUC, na.rm = TRUE),
    sd_ESUC = sd(ESUC, na.rm = TRUE),
    se_ESUC = sd_ESUC / sqrt(n),
    .groups = "drop"
  )

write.csv(age_summary, "tables/final/age_class_summary.csv", row.names = FALSE)

beta_age <- betareg(ESUC_prop_beta ~ AGECL, data = age_data)
beta_age_null <- betareg(ESUC_prop_beta ~ 1, data = age_data)
age_lrtest <- lmtest::lrtest(beta_age, beta_age_null)

age_emm <- emmeans(beta_age, ~ AGECL, type = "response")
age_pairs <- pairs(age_emm, adjust = "BH")

sink("outputs/model_summaries/beta_age_class.txt")
print(summary(beta_age))
print(age_lrtest)
print(age_emm)
print(age_pairs)
sink()

summary(beta_age)
age_lrtest
age_pairs
```

```{r figure-6-age-class, fig.width=10, fig.height=5}
fig6 <- ggplot(age_data, aes(x = AGECL, y = ESUC_prop)) +
  geom_boxplot(fill = "grey80", outlier.shape = NA, width = 0.6) +
  geom_jitter(width = 0.15, alpha = 0.35, size = 1.5) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Age class",
    y = "Emergence success"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig6
ggsave("figures/final/Figure_6_age_class_ESUC.png", fig6, width = 10, height = 5, dpi = 300)
```

```{r age-stage-models}
age_stage_data <- qmr_dataset2 %>%
  filter(!is.na(AGECL)) %>%
  mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
  filter(total_stage_dead > 0) %>%
  mutate(across(all_of(stage_cols), ~ .x / total_stage_dead, .names = "{.col}_prop")) %>%
  pivot_longer(
    cols = ends_with("_prop"),
    names_to = "stage",
    values_to = "prop"
  ) %>%
  mutate(
    stage = gsub("_prop", "", stage),
    stage_label = recode(stage, !!!stage_labels),
    stage_label = factor(stage_label, levels = stage_labels)
  )

age_stage_summary <- age_stage_data %>%
  group_by(AGECL, stage_label) %>%
  summarise(
    n = n(),
    mean_prop = mean(prop, na.rm = TRUE),
    se = sd(prop, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

write.csv(age_stage_summary, "tables/final/stage_mortality_by_age_class.csv", row.names = FALSE)
```

```{r figure-7-age-stage, fig.width=12, fig.height=6}
fig7 <- ggplot(age_stage_data, aes(x = AGECL, y = prop, fill = stage_label)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.75) +
  labs(
    x = "Age class",
    y = "Proportion of embryonic mortality",
    fill = "Developmental stage"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig7
ggsave("figures/final/Figure_7_stage_mortality_age_class.png", fig7, width = 12, height = 6, dpi = 300)
```

# 7. Temperature and rainfall

```{r temperature-rainfall-model, eval=TRUE}
# This section will run only if temperature and/or rainfall columns exist.
# If your temperature/rainfall file is separate, join it to qmr_dataset2 first
# and name the columns TEMPERATURE and RAINFALL.

weather_cols <- names(qmr_dataset2)[toupper(names(qmr_dataset2)) %in% c("TEMPERATURE", "TEMP", "TMAX", "RAINFALL", "RAIN")]
weather_cols

if ("TEMPERATURE" %in% names(qmr_dataset2)) {
  temp_data <- qmr_dataset2 %>%
    filter(!is.na(TEMPERATURE), !is.na(ESUC_prop_beta))

  beta_temp <- betareg(ESUC_prop_beta ~ TEMPERATURE, data = temp_data)
  beta_temp_null <- betareg(ESUC_prop_beta ~ 1, data = temp_data)

  if ("RAINFALL" %in% names(temp_data)) {
    beta_temp_rain <- betareg(ESUC_prop_beta ~ TEMPERATURE + RAINFALL, data = temp_data)
    temp_model_comparison <- AIC(beta_temp, beta_temp_rain)
  } else {
    beta_temp_rain <- NULL
    temp_model_comparison <- AIC(beta_temp, beta_temp_null)
  }

  sink("outputs/model_summaries/beta_temperature_rainfall.txt")
  print(summary(beta_temp))
  print(temp_model_comparison)
  sink()

  print(summary(beta_temp))
  print(temp_model_comparison)

  temp_pred <- data.frame(
    TEMPERATURE = seq(min(temp_data$TEMPERATURE, na.rm = TRUE),
                      max(temp_data$TEMPERATURE, na.rm = TRUE),
                      length.out = 100)
  )

  temp_pred$fit <- predict(beta_temp, newdata = temp_pred, type = "response")

  fig8 <- ggplot(temp_data, aes(x = TEMPERATURE, y = ESUC_prop)) +
    geom_point(alpha = 0.45) +
    geom_line(data = temp_pred, aes(x = TEMPERATURE, y = fit), linewidth = 1) +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
    labs(
      x = "Daily maximum temperature",
      y = "Emergence success"
    )

  print(fig8)
  ggsave("figures/final/Figure_8_temperature_beta_regression.png", fig8, width = 7, height = 5, dpi = 300)
} else {
  message("No TEMPERATURE column found in qmr_dataset2. Skipping temperature/rainfall model.")
}
```

# 8. Clutch number, relocation, and emergence success

```{r clutch-number-model}
clutch_data <- qmr_dataset2 %>%
  filter(!is.na(CNO), !is.na(Nest_management), CNO != 7) %>%
  mutate(
    CNO_factor = factor(CNO),
    TAG = factor(TAG)
  )

# Binomial GLMM using emergence success proportion with weights = 100.
# This follows the final approach used to test CNO × relocation status while accounting for repeated TAG.
glmm_clutch <- glmer(
  ESUC_prop ~ CNO * Nest_management + (1 | TAG),
  data = clutch_data,
  family = binomial(link = "logit"),
  weights = rep(100, nrow(clutch_data)),
  control = glmerControl(optimizer = "bobyqa")
)

glmm_clutch_null <- glmer(
  ESUC_prop ~ 1 + (1 | TAG),
  data = clutch_data,
  family = binomial(link = "logit"),
  weights = rep(100, nrow(clutch_data)),
  control = glmerControl(optimizer = "bobyqa")
)

glmm_clutch_no_interaction <- glmer(
  ESUC_prop ~ CNO + Nest_management + (1 | TAG),
  data = clutch_data,
  family = binomial(link = "logit"),
  weights = rep(100, nrow(clutch_data)),
  control = glmerControl(optimizer = "bobyqa")
)

clutch_lrt <- anova(glmm_clutch_no_interaction, glmm_clutch, test = "Chisq")

clutch_emm <- emmeans(glmm_clutch, ~ CNO | Nest_management, type = "response")
clutch_emm_table <- as.data.frame(clutch_emm)

write.csv(clutch_emm_table, "tables/final/clutch_number_emmeans.csv", row.names = FALSE)

sink("outputs/model_summaries/glmm_clutch_number_relocation.txt")
print(summary(glmm_clutch))
print(anova(glmm_clutch_null, glmm_clutch_no_interaction, glmm_clutch, test = "Chisq"))
print(clutch_lrt)
print(clutch_emm)
sink()

summary(glmm_clutch)
clutch_lrt
clutch_emm_table
```

```{r figure-9-clutch-number, fig.width=8, fig.height=5}
fig9_data <- clutch_data %>%
  group_by(CNO, Nest_management) %>%
  summarise(
    n = n(),
    mean_esuc = mean(ESUC_prop, na.rm = TRUE),
    se = sd(ESUC_prop, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

write.csv(fig9_data, "tables/final/clutch_number_observed_means.csv", row.names = FALSE)

fig9 <- ggplot(fig9_data, aes(x = CNO, y = mean_esuc, color = Nest_management, group = Nest_management)) +
  geom_point(size = 3) +
  geom_line(linewidth = 1) +
  geom_errorbar(aes(ymin = mean_esuc - se, ymax = mean_esuc + se), width = 0.1) +
  scale_x_continuous(breaks = sort(unique(fig9_data$CNO))) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Clutch number",
    y = "Mean emergence success ± SE",
    color = "Nest management"
  )

fig9
ggsave("figures/final/Figure_9_clutch_number_ESUC.png", fig9, width = 8, height = 5, dpi = 300)
```

```{r figure-10-proportion-relocated, fig.width=8, fig.height=5}
reloc_by_cno <- clutch_data %>%
  group_by(CNO) %>%
  summarise(
    n = n(),
    n_relocated = sum(Nest_management == "Relocated", na.rm = TRUE),
    prop_relocated = n_relocated / n,
    .groups = "drop"
  )

write.csv(reloc_by_cno, "tables/final/proportion_relocated_by_clutch_number.csv", row.names = FALSE)

fig10 <- reloc_by_cno %>%
  ggplot(aes(x = CNO, y = prop_relocated * 100)) +
  geom_col(alpha = 0.75, width = 0.6) +
  geom_text(aes(label = paste0("n = ", n)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_x_continuous(
    breaks = sort(unique(reloc_by_cno$CNO)),
    labels = paste0(sort(unique(reloc_by_cno$CNO)), c("st", "nd", "rd", rep("th", max(reloc_by_cno$CNO) - 3))[seq_along(sort(unique(reloc_by_cno$CNO)))])
  ) +
  ylim(0, max(reloc_by_cno$prop_relocated * 100, na.rm = TRUE) + 10) +
  labs(
    x = "Clutch number",
    y = "Clutches relocated (%)"
  )

fig10
ggsave("figures/final/Figure_10_proportion_relocated_by_clutch_number.png", fig10, width = 8, height = 5, dpi = 300)
```

# 9. Low-performing clutches

```{r low-performing-status}
low_summary <- badnest %>%
  group_by(Nest_management) %>%
  summarise(
    n = n(),
    mean_ESUC = mean(ESUC, na.rm = TRUE),
    sd_ESUC = sd(ESUC, na.rm = TRUE),
    se_ESUC = sd_ESUC / sqrt(n),
    .groups = "drop"
  )

write.csv(low_summary, "tables/final/low_performing_relocation_status_summary.csv", row.names = FALSE)

low_beta_reloc <- betareg(ESUC_prop_beta ~ Nest_management, data = badnest)
low_beta_reloc_null <- betareg(ESUC_prop_beta ~ 1, data = badnest)
low_reloc_lrtest <- lmtest::lrtest(low_beta_reloc, low_beta_reloc_null)

if (sum(!is.na(badnest$LDATE_numeric)) > 10) {
  low_beta_laydate <- betareg(ESUC_prop_beta ~ LDATE_numeric + Nest_management, data = badnest)
  low_laydate_compare <- AIC(low_beta_reloc, low_beta_laydate)
} else {
  low_beta_laydate <- NULL
  low_laydate_compare <- NA
}

sink("outputs/model_summaries/low_performing_relocation_status.txt")
print(summary(low_beta_reloc))
print(low_reloc_lrtest)
print(low_laydate_compare)
sink()

summary(low_beta_reloc)
low_reloc_lrtest
low_laydate_compare
```

```{r figure-11-low-status, fig.width=7, fig.height=5}
fig11 <- ggplot(badnest, aes(x = Nest_management, y = ESUC_prop, fill = Nest_management)) +
  geom_boxplot(alpha = 0.75, outlier.shape = NA, width = 0.55) +
  geom_jitter(width = 0.15, alpha = 0.45, size = 1.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Nest relocation status",
    y = "Emergence success",
    fill = "Nest management"
  ) +
  theme(legend.position = "none")

fig11
ggsave("figures/final/Figure_11_low_performing_relocation_status.png", fig11, width = 7, height = 5, dpi = 300)
```

```{r low-performing-stages}
low_stage_data <- badnest %>%
  mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
  filter(total_stage_dead > 0, !is.na(Nest_management)) %>%
  mutate(across(all_of(stage_cols), ~ .x / total_stage_dead, .names = "{.col}_prop")) %>%
  pivot_longer(
    cols = ends_with("_prop"),
    names_to = "stage",
    values_to = "prop"
  ) %>%
  mutate(
    stage = gsub("_prop", "", stage),
    stage_label = recode(stage, !!!stage_labels),
    stage_label = factor(stage_label, levels = stage_labels)
  )

low_stage_summary <- low_stage_data %>%
  group_by(Nest_management, stage_label) %>%
  summarise(
    n = n(),
    mean_prop = mean(prop, na.rm = TRUE),
    sd = sd(prop, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )

write.csv(low_stage_summary, "tables/final/low_performing_stage_mortality_summary.csv", row.names = FALSE)

low_stage_models <- map(stage_cols, function(stage_var) {
  data_stage <- badnest %>%
    mutate(total_stage_dead = rowSums(across(all_of(stage_cols)), na.rm = TRUE)) %>%
    filter(total_stage_dead > 0, !is.na(Nest_management)) %>%
    mutate(
      stage_prop = .data[[stage_var]] / total_stage_dead,
      stage_prop_beta = beta_transform(stage_prop)
    )

  model <- betareg(stage_prop_beta ~ Nest_management, data = data_stage)

  broom::tidy(model) %>%
    filter(component == "mean", term != "(Intercept)") %>%
    mutate(stage = stage_var)
}) %>%
  bind_rows() %>%
  mutate(
    p_holm = p.adjust(p.value, method = "holm"),
    stage_label = recode(stage, !!!stage_labels)
  )

write.csv(low_stage_models, "tables/final/low_performing_stage_beta_models.csv", row.names = FALSE)

low_stage_models
```

```{r figure-12-low-stage, fig.width=8, fig.height=5}
fig12 <- ggplot(low_stage_summary, aes(x = stage_label, y = mean_prop, color = Nest_management, group = Nest_management)) +
  geom_point(size = 3, position = position_dodge(width = 0.35)) +
  geom_line(position = position_dodge(width = 0.35)) +
  geom_errorbar(aes(ymin = mean_prop - se, ymax = mean_prop + se),
                width = 0.15, position = position_dodge(width = 0.35)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Developmental stage at death",
    y = "Mean proportion ± SE",
    color = "Nest management"
  ) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

fig12
ggsave("figures/final/Figure_12_low_performing_stage_mortality.png", fig12, width = 8, height = 5, dpi = 300)
```

```{r low-performing-relocation-type}
low_type_data <- badnest %>%
  filter(!is.na(RELOC_type))

low_kw_type <- kruskal.test(ESUC_prop ~ RELOC_type, data = low_type_data)

low_beta_type <- betareg(ESUC_prop_beta ~ RELOC_type, data = low_type_data)
low_beta_type_null <- betareg(ESUC_prop_beta ~ 1, data = low_type_data)
low_type_lrtest <- lmtest::lrtest(low_beta_type, low_beta_type_null)

low_type_emm <- emmeans(low_beta_type, ~ RELOC_type, type = "response")
low_type_pairs <- pairs(low_type_emm, adjust = "BH")

sink("outputs/model_summaries/low_performing_relocation_type.txt")
print(low_kw_type)
print(summary(low_beta_type))
print(low_type_lrtest)
print(low_type_emm)
print(low_type_pairs)
sink()

low_kw_type
summary(low_beta_type)
low_type_lrtest
low_type_pairs
```

```{r figure-13-low-type, fig.width=8, fig.height=5}
fig13 <- ggplot(low_type_data, aes(x = RELOC_type, y = ESUC_prop, fill = RELOC_type)) +
  geom_boxplot(alpha = 0.75, outlier.shape = NA, width = 0.55) +
  geom_jitter(width = 0.15, alpha = 0.45, size = 1.8) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Relocation type",
    y = "Emergence success",
    fill = "Relocation type"
  ) +
  theme(legend.position = "none")

fig13
ggsave("figures/final/Figure_13_low_performing_relocation_type.png", fig13, width = 8, height = 5, dpi = 300)
```

# 10. Additional non-parametric screening tests

```{r nonparametric-screens}
# These were useful screening tests but should only be reported if they remain in the thesis.

screen_tests <- list()

if ("SECT" %in% names(qmr_dataset2)) {
  screen_tests$ESUC_by_SECT <- kruskal.test(ESUC_prop ~ as.factor(SECT), data = qmr_dataset2)
}

if ("LDATE_numeric" %in% names(qmr_dataset2)) {
  # Kruskal-Wallis requires groups, so use LDATE as factor if desired.
  screen_tests$ESUC_by_LDATE <- kruskal.test(ESUC_prop ~ as.factor(LDATE_numeric), data = qmr_dataset2)
}

if ("RELOC" %in% names(qmr_dataset2)) {
  screen_tests$ESUC_by_RELOC_original <- kruskal.test(ESUC_prop ~ as.factor(RELOC), data = qmr_dataset2)
}

if ("TAG" %in% names(qmr_dataset2)) {
  # This can be statistically significant but hard to interpret because of many groups.
  screen_tests$ESUC_by_TAG <- kruskal.test(ESUC_prop ~ as.factor(TAG), data = qmr_dataset2)
}

sink("outputs/model_summaries/nonparametric_screening_tests.txt")
print(screen_tests)
sink()

screen_tests
```

# 11. Save final R session information

```{r session-info}
sink("outputs/model_summaries/session_info.txt")
sessionInfo()
sink()

sessionInfo()
```
